The Ky Fan Norms and Beyond: Dual Norms and Combinations for Matrix Optimization

### 2025-08-22 conf_article/neon.tex
Table for lp and Sp algorithms

@@

 Table that compares methods, updates, and sends to the proofs?

@@

\subsection{Algorithms for Matrices $\leftrightarrow$ Algorithms for Vectors}
Table that compares Muons and Neons to vector algorithms (NSGD).

>>>>

\usepackage{arydshln} % dashed lines

@@

@@

\subsection{Algorithms for Matrices $\leftrightarrow$ Algorithms for Vectors}

lmo optimizers in Schatten $S_p$ norms and in $l_p$ norms with common $p$ may be analogous to each other, as is illustrated by Table \ref{tbl:mat_vs_vec_lmo}. The analogies go beyond similarities in the updates. SignSGD is very close to Adam, as noted in \cite{bernstein2024oldoptimizernewnorm}, and both Adam and Muon perform well in training large models. NSGD is the same for both matrix and vector cases. Greedy Coordinate Descent methods are not applied to high-dimensional problems, from this perspective, it is not surprising that Neon underperforms on such problems.

Despite such similarities, no theory has been yet proposed that would reduce the matrix case to the optimization of a function of a singular numbers vector. If such a theory is developed, analysis of matrix algorithms like Muon will be greatly simplified.

\begin{table*}[t]
    \caption{lmo optimizers in Schatten $S_p$ norms and in $l_p$ norms. $g$ is the gradient. When it is a matrix, $g = \mU \mSigma \mV^\top$}
    \label{tbl:mat_vs_vec_lmo}
    \bgroup
    \def\arraystretch{1.2}
    \resizebox{\textwidth}{!}{
    \begin{tabular}{|l|c|c|c|}
    \hline
    Method & lmo constraint set $\mathcal D$ & lmo & Reference \\
    \hline
    \hline
    Normalized SGD & $l_2$-ball, $S_2$-ball & $-\eta \tfrac{g}{\norm{g}_2} = -\eta \tfrac{g}{\norm{g}_F}$ & \citep{hazan2015beyond} \\
    Momentum Normalized SGD & Ball in $l_2$, or Ball in $S_2$ & $-\eta \tfrac{g}{\norm{g}_2} = -\eta\tfrac{g}{\norm{g}_F}$ & \citep{cutkosky2020momentum}\\
    \hline
    SignSGD & Ball in Max-norm $l_\infty$ & $-\eta \sign(g)$ & \citep[Thm. 1]{bernstein2018signsgd} \\
    Signum  & Ball in Max-norm $l_\infty$ & $-\eta \sign(g)$ & \citep[Thm. 3]{bernstein2018signsgd} \\
    \hdashline
    Muon & Ball in Spectral $S_\infty$ & $-\eta UV^\top$ & \citep{jordan2024muon} \\
    \hline
    Gauss-Southwell Coordinate Descent & Ball in $l_1$ & $-\eta \{i: g_i \geq g_k \forall k\}$ & \citep[p. 19]{shi2016primer}\\
    \hdashline 
    Neon & Ball in Nuclear $S_1$ & $-\eta u_1 v_1^\top$ & This work\\
    \hline
    \end{tabular}
    }
    \egroup
    \end{table*}
